\textbf{Emergency Service Location Optimization in a Multi-Zone County based on Population Distribution}

### 2013-11-16 LaTeX/paper.tex
formatting
further progress on ambulance distribution

\title{\textbf{Emergency Service Location Optimization in a Multi-Zone District based on Population Distribution}}
\author{Team 4552}

@@

\tableofcontents
\newpage
\section{Problem Restatement}
Given a district made up of six separate zones and average travel times between zones, create a model to optimize the placement of $n$ ambulances in order to maximize the number of civilians reached in an eight minute period. Consider the scenarios where $n = 3$, $n = 2$, and $n = 1$, respectively, and for each scenario, keep track of how many people are not being reached with each possible solution.

Then, consider a scenario in which a large scale disaster affects a single location (e.g. the September 11th attacks), and discuss how an Emergency Service Coordinator would cover the situation. Examine how a real-world city or district would prepare for such a disaster. Finally, write a two page memo detailing the model and its analysis for the Emergency Service Coordinator.

\section{Assumptions and Justifications}
\begin{description}
    \item[TODO(drew): Add these assumptions] \hfill \\
    	* Each zone is not a point, but rather a distribution. We experimented with...
        * Route frequency is correlated with population frequency (two random variables?)
    \item[1. Naming] \hfill \\
    	A district is another name for a county. ... TODO(drew)
        Travel...
    \item[2. Zone to Zone Travel] \hfill \\
    	The travel times given are considered to be optimal routes from the center of one zone to the center of the another, regardless of any other zones that may lie in between.
    \item[3. Travel Inside Each Zone] \hfill \\
    	The travel times given for travel from any zone to itself is the average
        of all the travel times of the different routes within that zone.
    \item[4. Directional Travel Inside Each Zone] \hfill \\
    	The ratio between the time it takes inside a zone to travel toward a second zone,
        to the time it takes inside a zone to return from a second, is equal to the ratio
        between the travel time from the first zone to the second, and the travel time from the
        second to the first.
    \item[5. Ambulance Placement] \hfill \\
        Ambulances can only be placed in the very center of a zone. Because we do not know the
        spacial position of the zones relative to each other, we cannot consider the ambulances
        being closer to one zone or another.
    \item[6. Partial Coverage] \hfill \\
        Partial ambulance coverage of a zone will be treated as incomplete even coverage over the entire
        zone. Because ambulances of ideal placement are unlikely to overlap in partial coverage, we assume
        that multiple ambulances will cover different areas of the zones, thus evening out coverage. (If partial coverages from multiple ambulances sum to above the total population of a zone,
        we assume that zone is completely covered by the ambulances.)
\end{description}

@@

Therefore, a more scalable solution is needed. We discovered that a greedy heuristic also gives a good solution,
except with a much lower time complexity of $O(n, k) = n k^2$.

@@

\subsection{Population Distribution}
As it currently stands, we have very little information regarding the actual population distribution of each zone in the scope of the district, which provides us with a slightly skewed perception. While we justified the use of a standard distribution for population above, we have the ability to greatly improve our model with a better idea of how spread out a population of a zone actually is, in relation to its epicenter. This information allows us to better predict how many people an ambulance would miss (not be able to cover) in a certain zone.

@@

\item Identify a set of nodes such that there is a large number of outgoing edges, thereby identifying potential split points, to divide each district into smaller subdistricts. Make sure to account for overlap.

\item Allocate $n$ ambulances across the subdistricts. If there are fewer ambulances than are necessary to cover the entire district, weight each subdistrict by population size (save more people per ambulance).

\item Solve each subdistrict for the optimal ambulance location, thereby finding a solution for the set as a whole.
\end{enumerate}

>>>>

\title{\textbf{Emergency Service Location Optimization in a Multi-Zone County based on Population Distribution}}
\author{Team 4552}

@@

\tableofcontents
\newpage
\section{Problem Restatement}
Given a county made up of six separate zones and average travel times between zones, create a model
to optimize the placement of $n$ ambulances in order to maximize the number of civilians reached in an
eight minute period. Consider the scenarios where $n = 3$, $n = 2$, and $n = 1$, respectively, and for
each scenario, keep track of how many people are not being reached with each possible solution.

Then, consider a scenario in which a large scale disaster affects a single location (e.g. the September 11th attacks),
and discuss how an Emergency Service Coordinator would cover the situation. Examine how a real-world city or
county would prepare for such a disaster. Finally, write a two page memo detailing the model and its analysis
for the Emergency Service Coordinator.

\section{Assumptions and Justifications}
\begin{description}
    \item[TODO(drew): Add these assumptions] \hfill \\
    	* Each zone is not a point, but rather a distribution. We experimented with...
        * Route frequency is correlated with population frequency (two random variables?)
    \item[1. Naming] \hfill \\
    	A county is another name for a county. ... TODO(drew)
        Travel...
    \item[2. Zone to Zone Travel] \hfill \\
    	The travel times given are considered to be optimal routes from the center of one zone to
        the center of the another, regardless of any other zones that may lie in between.
    \item[3. Travel Inside Each Zone] \hfill \\
    	The travel times given for travel from any zone to itself is the average
        of all the travel times of the different routes within that zone.
    \item[4. Directional Travel Inside Each Zone] \hfill \\
    	The ratio between the time it takes inside a zone to travel toward a second zone,
        to the time it takes inside a zone to return from a second, is equal to the ratio
        between the travel time from the first zone to the second, and the travel time from the
        second to the first.
    \item[5. Ambulance Placement] \hfill \\
        Ambulances can only be placed in the very center of a zone. Because we do not know the
        spacial position of the zones relative to each other, we cannot consider the ambulances
        being closer to one zone or another.
    \item[6. Partial Coverage] \hfill \\
        Partial ambulance coverage of a zone will be treated as incomplete even coverage over the entire
        zone. Because ambulances of ideal placement are unlikely to overlap in partial coverage, we assume
        that multiple ambulances will cover different areas of the zones, thus evening out coverage.
        (If partial coverages from multiple ambulances sum to above the total population of a zone,
        we assume that zone is completely covered by the ambulances.)
\end{description}

@@

Therefore, a more scalable solution is needed. We discovered that a greedy heuristic also gives a good solution,
except with a much lower time complexity of $O(n, k) = n k^2$. The greedy heuristic works by assigning ambulances to zones
one at a time. For each assignment, we iterate through each zone without an ambulance already assigned, and see how many
additional people it would cover; then we assign an ambulance to the zone which would most increase the number of
covered people.

@@

\subsection{Population Distribution}
As it currently stands, we have very little information regarding the actual population distribution of each
zone in the scope of the county, which provides us with a slightly skewed perception. While we justified the
use of a standard distribution for population above, we have the ability to greatly improve our model with a
better idea of how spread out a population of a zone actually is, in relation to its epicenter. This information
allows us to better predict how many people an ambulance would miss (not be able to cover) in a certain zone.

@@

\item Identify a set of nodes such that there is a large number of outgoing edges, thereby identifying potential
    split points, to divide each county into smaller subcounties. Make sure to account for overlap.

\item Allocate $n$ ambulances across the subcounties. If there are fewer ambulances than are necessary to cover
    the entire county, weight each subcounty by population size (save more people per ambulance).

\item Solve each subcounty for the optimal ambulance location, thereby finding a solution for the set as a whole.
\end{enumerate}
